Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-21 paper/paper.tex
closer, in better shape

The bolometric UV luminosity is $9.43\pm1.94 \times 10^{8}$
L$_{\odot}$ as measured by GALEX, without any correction by extinction
and following the empirical relation by \cite{Kennicutt98}, this
corresponds to a star formation rate of $0.35\pm 0.05$ M$_{\odot}$
yr$^{-1}$. 
Its metallicity is $\sim Z_{\odot}/24$ as derived from
optical spectroscopy \citep{Izotov04}.

%The absolute magnitude in the $V$ band translates into a luminosity of
%$8.9\times 10^{8}$ L$_{\odot}$.  
% using http://tomdwelly.com/tools_fluxtolum.php
%

@@

\tol\ shows one of the most atypical \lya\ spectra observed so far.
Its flux at the line's center is high compared to other LAEs at low
redshift and the wide symmetrical peak is virtually absent from other
LAEs at low and high redshift \citep{2012ApJ...751...29Y,LARS,Erb14,Trainor16}. Simple
shell models fail to reproduce such a spectrum
\cite{2015A&A...578A...7V}.  
In the previous sections we show how these characteristics can be
explained by rather extreme kinematic characteristics.
Now we discuss and frame those results in terms of the dynamical mass
that could explain the velocities obtained in the model.

Using the constraints for the velocity dispersion $\sigma$ and the
maximum rotational velocity $V_m$ in a spherical system localized in a
region of size $r$ we can estimate two different values for the
dynamical mass within $r$:

\begin{equation}
M_{\rm dyn} = 3 \frac{\sigma ^{2}r}{G} = 3.48\times10^{9}
\left(\frac{\sigma}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)
M_{\odot}, 
\end{equation}
%
and
%
\begin{equation}
M_{\rm dyn} = \frac{V_m ^{2}r}{G} = 1.16\times10^{9}
\left(\frac{V_m}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)M_{\odot}, 
\end{equation}
%
for the multiphase and rotation model, respectively.
Assuming that the \lya\ emission is entirely powered by star formation 
we use the 3D half-light radius $r_s=2.25$ kpc as the typical size
for the HI region,

In the multiphase model the clump velocity dispersion
$\sigma_{\rm{cl}}=$\sigmaclump corresponds to a dynamical mass of
$M_{\rm dyn}=2.31\pm0.04 \times 10^{9}$ $M_{\odot}$, while the
rotational velocity of $V_{\rm max}=348^{+75}_{-48}$\kms corresponds
$M_{\rm dyn} = 3.2_{-1.0}^{+1.6}\times 10^{10} M_{\odot}$.

\tol's stellar mass is  $M_{\star} = 1.45\pm0.45\times 10^{8}
M_{\odot}$   \citep{2014PASP..126.1079M} and its total neutral HI mass
is $M_{\rm HI}<2.65\times 10^{8}$ M$_{\odot}$ 
\citep{pustilnikmartin07}; the dynamical mass is at least 12 to 160 times
the baryonic mass, depending if one considers the multiphase or
bulk rotation estimate. 
We lean towards the lower dynamical mass estimate from the multiphase
model as it seems easier to reconcile with the following two
astrophysical mechanisms for its origin.

A first possibility to explain a dynamical mass of $10^{9} M_{\odot}$
in a sphere of 2.25 kpc in radius, could be a dark matter halo of at
least $10^{12}$ M$_{\odot}$ in mass \citep{2011ApJ...726..108T}.
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the \emph{median} mass-concentration
relationship found in the Bolshoi simulation
\citep{2012MNRAS.423.3018P}. 
This estimate leaves open the question as to why  \tol\ is not more
similar to the Milky Way galaxy as it would be hosted by a dark matter
halo of similar mass.

A second possibility is that \tol\ hosts a supermassive black hole of
$10^{9} M_{\odot}$. This is almost two orders of magnitude higher
than the supermassive black hole found in the compact dwarf galaxy
M60-UCD1 \citep{2014Natur.513..398S}, which has a similar stellar mass
as \tol. This would leave open the question about the formation
process of such a system.

%Another perspective to appreciate the atypically high dynamical mass
%estimates comes from the observed scaling relations for dwarf
%galaxies.
%Assuming that \tol\ followed the fundamental plane relationship
%between its mean surface brightness $I_e$, the projected half-light
%radius $R_e$ and the velocity dispersion $\sigma$, described by $\log
%I_e=1.6 \log\sigma - 1.21\log R_e + 0.55$ \citep{2009ApJ...698.1590G},
%the expected velocity dispersion should be on the order of $5 \pm 1$ 
%\kms, which is a factor of $\sim$ $10$ - $60$ lower than the 
%results from the multiphase and rotation models, respectively. 
%These are equivalent to factors of $\sim$ $100$ - $3600$ on the
%dynamical mass.
%Once again,  \tol\ seems to be significantly more massive than
%expected.

A new observational test is needed to clarify the kinematic nature of
\tol. 
We suggest that integral field unit measurements 
resolving its spatial extent are up to the task. 
\tol\ spans a region of $4$ arcseconds,
an instrument such as the Multi Unit Spectroscopic Explorer (MUSE)
\citep{2014Msngr.157...13B} with its
nominal $0.2$ arcseconds spatial sampling over a $1.0$ arcminute field
in wide-field mode could provide a coarse mapping of different
ionization lines to infer a kinematic map in the H$\alpha$ line,
following the kind of work presented by \citet{Herenz16} done with the Potsdam
Multi-Aperture Spectrometer (PMAS) \citet{PMAS} on nearby
\lya\ emitting galaxies.

@@

We acknowledge that a final confirmation of our kinematic findings has
to be provided by the new observational tests we suggest.
If that were the case, this would open up a new line of more detailed
enquiries to understand how a compact dwarf galaxy can have such a
large dynamical mass.

>>>>

The bolometric UV luminosity is $9.43\pm1.94 \times 10^{8}$
L$_{\odot}$ as measured by GALEX, without any correction by extinction
and following the empirical relation by \cite{Kennicutt98}, this
corresponds to a star formation rate of $0.35\pm 0.05$ M$_{\odot}$
yr$^{-1}$. 
The absolute magnitude in the $V$ band translates into a luminosity of
$8.9\times 10^{8}$ L$_{\odot}$.  
Its metallicity is $\sim Z_{\odot}/24$ as derived from
optical spectroscopy \citep{Izotov04}. 
% using http://tomdwelly.com/tools_fluxtolum.php
%

@@

\tol\ shows one of the most atypical \lya\ spectra observed so far.
Its flux at the line's center is high compared to other LAEs at low
redshift and the wide symmetrical peak is virtually absent from other
LAEs at low and high redshift \citep{2012ApJ...751...29Y,LARS,Erb14,Trainor16}. Simple
shell models fail to reproduce such a spectrum
\cite{2015A&A...578A...7V}.  
In the previous sections we show how these characteristics can be
explained by rather extreme kinematic characteristics.

A first perspective to put in context \tol's kinematic is comes the
observed scaling relations for elliptic galaxies.
Assuming that \tol\ follows the fundamental plane relationship
between its mean surface brightness $I_e=L_V/2\pi R_{e}^2$, the
projected half-light radius $R_e$ and the velocity dispersion
$\sigma$, described by $\log I_e=1.6 \log\sigma - 1.21\log R_e + 0.55$ \citep{2009ApJ...698.1590G},
the expected velocity dispersion should be on the order of $5\pm 1$
\kms, which is a factor of $\sim\ 10$ - $60$ lower than the   
results from the multiphase and rotation models, respectively.

However, \tol's structural properties, seem to be closer to observed dwarf
elliptical galaxies \citep{2003AJ....126.1794G}. 
Galaxies such as IC 3344 and NGC 4482 have similar absolute $M_V$
magnitudes  and size as \tol. 
These galaxies have velocity dispersions in the range $30$-$45$ \kms. 
The rotational velocities of those systems is lower than $1$\kms. 
This gives support to the idea that the multiphase model might be
closer to be the correct model describing \tol, as it provides at
least kinematic consistency with other dwarf galaxy populations.

Using the constraints for the velocity dispersion $\sigma$  in a
spherical system localized in a region of size $r$ we can estimate a
value or the dynamical mass within $r$:

\begin{equation}
M_{\rm dyn} = 3 \frac{\sigma ^{2}r}{G} = 3.48\times10^{9}
\left(\frac{\sigma}{100\ \mathrm{km\ s}^{-1}}\right)^2\left(\frac{r}{\mathrm{kpc}}\right)
M_{\odot}, 
\end{equation}

Assuming that the \lya\ emission is entirely powered by star formation 
we use the 3D half-light radius $r_s=2.25$ kpc as the typical size
for the HI region, in the multiphase model the clump velocity dispersion
$\sigma_{\rm{cl}}=$\sigmaclump corresponds to a dynamical mass of
$M_{\rm dyn}=2.31\pm0.04 \times 10^{9}$ $M_{\odot}$.

Following the results of \cite{2011ApJ...726..108T}, who made a
detailed analysis for dwarf ellipticals, we have that such dynamical
mass can be explained if \tol\ is hosted by a dark matter halo of at
least $10^{11}$ M$_{\odot}$ in mass. 
This estimate is based on computing the integrated mass profile of a
spherical dark matter halo with a Navarro-Frenk-White profile with its
concentration following the \emph{median} mass-concentration
relationship found in the Bolshoi simulation \citep{2012MNRAS.423.3018P}.

This results would imply that \tol\ is part of the dwarf ellipticals
class. However, there are two problems that remain. First, dwarf
ellipticals are found either in clusters or groups of galaxies; while
\tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F,}. The second problem is that, unlike dwarf
ellipticals, \tol\ hosts bright an compact startburst; imaging and
spectroscopy studies suggest that the startburst started less than 4
Myr ago, while the low surface brightness regins are consistent with
an age below one Gyr \citep{2001AJ....121..169F}.

New observational tests are needed to clarify the kinematic nature of
\tol. 
We suggest that integral field unit measurements 
resolving its spatial extent are up to the task. 
\tol\ spans a region of $4$ arcseconds,
an instrument such as the Multi Unit Spectroscopic Explorer (MUSE)
\citep{2014Msngr.157...13B} with its
nominal $0.2$ arcseconds spatial sampling over a $1.0$ arcminute field
in wide-field mode could provide a coarse mapping of different
ionization lines to infer a kinematic map in the H$\alpha$ line,
following the kind of work presented by \citet{Herenz16} done with the Potsdam
Multi-Aperture Spectrometer (PMAS) \citet{PMAS} on nearby
\lya\ emitting galaxies.

@@

We acknowledge that a final confirmation of our kinematic findings has
to be provided by new observational tests of the kind we suggest.
If that were the case, this would open up a new line of more detailed
enquiries to understand how a compact dwarf galaxy can have such a
large dynamical mass.
